---
title: "Project"
author: "Sri Harsha Sudalagunta"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lets load the datasets in to the file

Diabetes Datasets

```{r}
diabetes_data <- read.csv("C:/Users/SRI HARSHA S/OneDrive - Indiana University/Documents/R/Project Files/diabetes_data.csv")
print(diabetes_data)


```

Lets See the description of the data

```{r}

library(Hmisc)
describe(diabetes_data)
```

We can see here that there are no missing values in the data set and we can see that many of them are categorical.

Lets see whether we have enough data of the samples who are diabetic

```{r}
sum(diabetes_data$Diabetes)/nrow(diabetes_data)
```

Lets see the visualizations of some of the dataset install.packages("ggplot2")

```{r}
# create a table with counts of diabetes cases by age group
age_table <- table(diabetes_data$Diabetes, diabetes_data$Age)

# create a bar chart
barplot(age_table, main = "Diabetes Cases by Age Group", xlab = "Age Group", ylab = "Number of Cases")



```

Sex and diabetes

```{r}
# create a table with counts of diabetes cases by age group
age_table <- table(diabetes_data$Diabetes, diabetes_data$Sex)

# create a bar chart
barplot(age_table, main = "Diabetes Cases by Age Group", xlab = "Sex", ylab = "Number of Cases")

```

Lets See our 1st Research Question:

```{r}
table(diabetes_data$Age, diabetes_data$Diabetes)
chisq.test(diabetes_data$Age, diabetes_data$Diabetes)

table(diabetes_data$Sex, diabetes_data$Diabetes)
chisq.test(diabetes_data$Sex, diabetes_data$Diabetes)

table(diabetes_data$HighChol, diabetes_data$Diabetes)
chisq.test(diabetes_data$HighChol, diabetes_data$Diabetes)

table(diabetes_data$CholCheck, diabetes_data$Diabetes)
chisq.test(diabetes_data$CholCheck, diabetes_data$Diabetes)

table(diabetes_data$Smoker, diabetes_data$Diabetes)
chisq.test(diabetes_data$Smoker, diabetes_data$Diabetes)

table(diabetes_data$PhysActivity, diabetes_data$Diabetes)
chisq.test(diabetes_data$PhysActivity, diabetes_data$Diabetes)

table(diabetes_data$HvyAlcoholConsump, diabetes_data$Diabetes)
chisq.test(diabetes_data$HvyAlcoholConsump, diabetes_data$Diabetes)

table(diabetes_data$GenHlth, diabetes_data$Diabetes)
chisq.test(diabetes_data$GenHlth, diabetes_data$Diabetes)

table(diabetes_data$MentHlth, diabetes_data$Diabetes)
t.test(diabetes_data$MentHlth ~ diabetes_data$Diabetes)

table(diabetes_data$PhysHlth, diabetes_data$Diabetes)
t.test(diabetes_data$PhysHlth ~ diabetes_data$Diabetes)

```

Interpretation: The results of the Chi-squared tests indicate a significant association between all the predictor variables (Age, Sex, HighChol, CholCheck, Smoker, PhysActivity, HvyAlcoholConsump, and GenHlth) and the response variable (Diabetes), with p-values less than 0.05.
This suggests that each of these predictor variables is associated with the risk of developing diabetes.

The Welch Two Sample t-test also indicates a significant difference in the mean values of the Mental Health variable between those with diabetes and those without diabetes, with a p-value less than 0.05.
This suggests that there is a significant difference in mental health between those with diabetes and those without diabetes.

Lets do the logistic regression

```{r}
# assuming your data frame is named "data" with columns "age", "sex", and "diabetes"
logit_model <- glm(Diabetes ~ Age + Sex + HighChol + CholCheck + Smoker + PhysActivity + 
                   HvyAlcoholConsump + GenHlth + MentHlth + PhysHlth, data = diabetes_data, 
                   family = binomial(link = "logit"))
summary(logit_model)


```

Interpretation

Based on the logistic regression model output, we can make the following findings:

Age is positively associated with the log-odds of having diabetes.
For every one-unit increase in age, the log-odds of having diabetes increase by 0.165 (or the odds increase by exp(0.165) = 1.18).
Females are more likely to have diabetes than males, as indicated by the positive coefficient for Sex.
High cholesterol levels are positively associated with the log-odds of having diabetes.
For every one-unit increase in HighChol, the log-odds of having diabetes increase by 0.780 (or the odds increase by exp(0.780) = 2.18).
Having had a cholesterol check is also positively associated with the log-odds of having diabetes, with a slightly smaller effect size than HighChol.
Engaging in physical activity is negatively associated with the log-odds of having diabetes.
For every one-unit increase in PhysActivity, the log-odds of having diabetes decrease by 0.230 (or the odds decrease by exp(-0.230) = 0.80).
Heavy alcohol consumption is negatively associated with the log-odds of having diabetes, indicating that individuals who consume heavy amounts of alcohol are less likely to have diabetes.
Poor general health (GenHlth) is positively associated with the log-odds of having diabetes.
Mental health (MentHlth) is not a significant predictor of diabetes, as indicated by its non-significant p-value.
Physical health (PhysHlth) is negatively associated with the log-odds of having diabetes.
For every one-unit increase in PhysHlth, the log-odds of having diabetes decrease by 0.005 (or the odds decrease by exp(-0.005) = 0.99).
Overall, the findings suggest that age, sex, cholesterol levels, physical activity, heavy alcohol consumption, general health, and physical health are all important predictors of diabetes.

```{r}
install.packages("metafor")
# Load required package
library(metafor)

# Create a dataframe with the coefficients and confidence intervals
coef_df <- data.frame(
  coef = coef(logit_model),
  ci_low = confint(logit_model)[,1],
  ci_high = confint(logit_model)[,2]
)

# Add a column with variable names
coef_df$var <- row.names(coef_df)

# Create a forest plot
forest(coef_df$coef, ci.lb = coef_df$ci_low, ci.ub = coef_df$ci_high,
       slab = coef_df$var, xlim = c(-6, 2), ilab = cbind(coef_df$coef, coef_df$var))

```

2.  Are there any significant differences in the risk of developing diabetes between different demographic groups?

```{r}

# Load the diabetes data
diabetes_data <- read.csv("diabetes_data.csv")

# Calculate the number of individuals with diabetes by sex
diabetes_by_sex <- table(diabetes_data$Sex, diabetes_data$Diabetes)

# Calculate the prevalence of diabetes by sex
prevalence_by_sex <- diabetes_by_sex["1","0"]/ rowSums(diabetes_by_sex)

# Print the prevalence by sex
prevalence_by_sex

```

Chi square test for demographic groups age and sex and diabetes.

```{r}
# Create contingency table
table <- table(diabetes_data$Diabetes, diabetes_data$Age, diabetes_data$Sex)

# Perform chi-square test
(table)

```

```{r}
# Perform chi-square test
# Convert to 2-dimensional table
table_2d <- table[, , 1] + table[, , 2]

# Perform chi-square test
chisq.test(table_2d)

```

```{r}
# create a contingency table
Agetable <- table(diabetes_data$Age, diabetes_data$Diabetes)

# perform the chi-square test
chisq.test(Agetable)

Sextable <- table(diabetes_data$Sex, diabetes_data$Diabetes)

# perform the chi-square test
chisq.test(Sextable)


```

Calculate the mean BMI for individuals with and without diabetes, and compare them.
You can use a t-test to determine whether the difference in means is statistically significant.

```{r}
# Subset the data into two groups: individuals with and without diabetes
diabetes <- subset(diabetes_data, Diabetes == "1")
no_diabetes <- subset(diabetes_data, Diabetes == "0")

# Calculate the mean BMI for each group
mean_bmi_diabetes <- mean(diabetes$BMI)
mean_bmi_no_diabetes <- mean(no_diabetes$BMI)

# Perform a t-test to determine whether the difference in means is significant
t.test(diabetes$BMI, no_diabetes$BMI)

```

Create a boxplot of BMI for individuals with and without diabetes to visualize any differences in distribution.

```{r}
# Create a boxplot of BMI for each group
boxplot(BMI ~ Diabetes, data = diabetes_data)

```

To determine whether there is a significant association between BMI and diabetes prevalence, you can perform a chi-squared test of independence.
This test will assess whether there is a significant difference in the proportion of individuals with diabetes across different categories of BMI.

```{r}
# Perform a chi-squared test of independence to determine whether there is a significant association between BMI and diabetes prevalence
bmitable <- table(diabetes_data$BMI, diabetes_data$Diabetes)
bmitable
chisq.test(bmitable)
```

Research question 3:

What are the most common medical conditions thatco occur with diabetes and are there any significant associations between tese conditions and the diabetes status?

```{r}
table(diabetes_data$HeartDiseaseorAttack, diabetes_data$Diabetes)
chisq.test(diabetes_data$HeartDiseaseorAttack, diabetes_data$Diabetes)

table(diabetes_data$Stroke, diabetes_data$Diabetes)
chisq.test(diabetes_data$Stroke, diabetes_data$Diabetes)

table(diabetes_data$HighBP, diabetes_data$Diabetes)
chisq.test(diabetes_data$HighBP, diabetes_data$Diabetes)


```
